---
title: "Introdruction to courseraFarsR"
output: rmarkdown::html_vignette
description: >
  This vignette should grant you a smooth start into
  handling FARS accident data with the courseraFarsR
  package.
vignette: >
  %\VignetteIndexEntry{courseraFarsR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Disclaimer
*This is a purely educational and private project. Just set to public to interact with [travis ci](https://travis-ci.org) and give those peer reviewers at coursera a chance to peek.*

## Overview
The package at hand provides you with some handy functions for handling data delivered by the
[Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars) (FARS). The main purpose of the package is to import the FARS data in a tidy format and help the user plotting and summarizing the data for certain years. The key function is `fars_map_state()` which enables you to map accident data for a given state number and year.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`fars_read()` provides you with a reader function for the compress FARS csv data. It produces a tidy tibble, as an example the 2013 data (`accident_2013.csv.bz2`) have been read in:

### Importing Data: accident_2013.csv.bz2
```{r read_in}
library(courseraFarsR)
fars_read(system.file("extdata", "accident_2013.csv.bz2", package = "courseraFarsR"))
```

### Helpers: construct filename and read multiple years
The respective filename for a certain year's data according to the FARS standard can be generated with the function `make_filename()`:
```{r make_filename}
make_filename(2013)
```

For the purpose of summarizing the information later on, multiple years can be imported in the following way as month-year combinations. The function returns a list object each element being a tibble of month:
```{r fars_read_years}
fars_read_years(c(2013, 2015))
```

### Compute summary and map data
Those combinations can also be directly summarized with the help of the `make_filename()` function. This gives you the possibility to compare the monthly accident counts of multiple years:
```{r fars_summarize_years}
fars_summarize_years(c(2013, 2015))
```

The package's main function `fars_map_state()` makes use of the helper functions above map the accident data for a given state number and a given year:
```{r plot}
fars_map_state(1, 2014)
```

